---
title: "MaxDiff Analysis of MSBA Class Preferences"
subtitle: "Counts, MLE, and Bayesian estimation of the MNL model"
author: "Qiang Hu"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    theme: cosmo
    code-fold: true
    code-tools: true
    embed-resources: true
    html-math-method: mathjax
    fig-format: png
    fig-align: center
execute:
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(knitr)
library(scales)

set.seed(495)

# Small helper for cleaner tables
fmt_pct <- function(x) scales::percent(x, accuracy = 0.1)
fmt_num <- function(x, digits = 3) round(x, digits)


# Inline SVG helpers avoid external figure files, so the blog page renders cleanly
# even when the site is previewed, published, or saved as a PDF.
escape_html <- function(x) {
  x <- as.character(x)
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  x <- gsub(">", "&gt;", x, fixed = TRUE)
  x <- gsub('"', "&quot;", x, fixed = TRUE)
  x
}

svg_bar_chart <- function(data, label, value, title, low = NULL, high = NULL,
                          percent = FALSE, width = 820, row_height = 34,
                          margin_left = 330, margin_right = 95) {
  d <- data |>
    dplyr::mutate(.label = .data[[label]], .value = .data[[value]]) |>
    dplyr::arrange(dplyr::desc(.value))

  if (!is.null(low)) d$.low <- d[[low]] else d$.low <- d$.value
  if (!is.null(high)) d$.high <- d[[high]] else d$.high <- d$.value

  xmin <- min(c(d$.low, d$.value, 0), na.rm = TRUE)
  xmax <- max(c(d$.high, d$.value, 0), na.rm = TRUE)
  if (xmax == xmin) xmax <- xmin + 1

  plot_w <- width - margin_left - margin_right
  top <- 56
  height <- top + nrow(d) * row_height + 22
  scale_x <- function(x) margin_left + (x - xmin) / (xmax - xmin) * plot_w
  zero_x <- scale_x(0)
  fmt_val <- function(x) if (percent) scales::percent(x, accuracy = 0.1) else sprintf("%.3f", x)

  rows <- purrr::map_chr(seq_len(nrow(d)), function(i) {
    y <- top + (i - 1) * row_height
    x0 <- scale_x(0)
    xv <- scale_x(d$.value[i])
    rect_x <- min(x0, xv)
    rect_w <- max(abs(xv - x0), 1)
    ci <- ""
    if (!is.null(low) && !is.null(high)) {
      xl <- scale_x(d$.low[i]); xh <- scale_x(d$.high[i]); cy <- y + 13
      ci <- paste0(
        "<line x1='", xl, "' x2='", xh, "' y1='", cy, "' y2='", cy, "' stroke='#1f2937' stroke-width='1.2'/>",
        "<line x1='", xl, "' x2='", xl, "' y1='", cy - 5, "' y2='", cy + 5, "' stroke='#1f2937' stroke-width='1.2'/>",
        "<line x1='", xh, "' x2='", xh, "' y1='", cy - 5, "' y2='", cy + 5, "' stroke='#1f2937' stroke-width='1.2'/>"
      )
    }
    paste0(
      "<text x='12' y='", y + 17, "' font-size='12' fill='#111827'>", escape_html(d$.label[i]), "</text>",
      "<rect x='", rect_x, "' y='", y, "' width='", rect_w, "' height='22' rx='4' fill='#2563eb' opacity='0.82'/>",
      ci,
      "<text x='", width - margin_right + 12, "' y='", y + 16, "' font-size='12' fill='#111827'>", fmt_val(d$.value[i]), "</text>"
    )
  })

  paste0(
    "<div style='margin:1rem 0 1.25rem 0; overflow-x:auto;'>",
    "<svg viewBox='0 0 ", width, " ", height, "' width='100%' style='max-width:", width, "px; font-family:system-ui,-apple-system,Segoe UI,sans-serif;'>",
    "<text x='12' y='28' font-size='17' font-weight='700' fill='#111827'>", escape_html(title), "</text>",
    "<line x1='", zero_x, "' x2='", zero_x, "' y1='42' y2='", height - 12, "' stroke='#6b7280' stroke-width='1'/>",
    paste(rows, collapse = ""),
    "</svg></div>"
  )
}

svg_trace_plot <- function(post_beta, burn) {
  trace_df <- as_tibble(post_beta) |>
    mutate(iteration = (burn + 1):20000) |>
    select(iteration, beta_2, beta_5, beta_9) |>
    pivot_longer(-iteration, names_to = "parameter", values_to = "value")

  width <- 820; panel_h <- 145; left <- 60; right <- 20; top <- 42
  height <- top + 3 * panel_h + 20

  panels <- purrr::map_chr(unique(trace_df$parameter), function(par) {
    idx <- match(par, unique(trace_df$parameter))
    d <- trace_df |> filter(parameter == par) |> slice(seq(1, n(), by = 20))
    xmin <- min(d$iteration); xmax <- max(d$iteration)
    ymin <- min(d$value); ymax <- max(d$value)
    if (ymax == ymin) ymax <- ymin + 1
    y0 <- top + (idx - 1) * panel_h
    sx <- function(x) left + (x - xmin) / (xmax - xmin) * (width - left - right)
    sy <- function(y) y0 + panel_h - 28 - (y - ymin) / (ymax - ymin) * (panel_h - 48)
    pts <- paste0(round(sx(d$iteration), 1), ",", round(sy(d$value), 1), collapse = " ")
    paste0(
      "<text x='12' y='", y0 + 18, "' font-size='12' font-weight='700' fill='#111827'>", par, "</text>",
      "<line x1='", left, "' x2='", width - right, "' y1='", y0 + panel_h - 28, "' y2='", y0 + panel_h - 28, "' stroke='#d1d5db'/>",
      "<polyline points='", pts, "' fill='none' stroke='#2563eb' stroke-width='1.2' opacity='0.9'/>"
    )
  })

  paste0(
    "<div style='margin:1rem 0 1.25rem 0; overflow-x:auto;'>",
    "<svg viewBox='0 0 ", width, " ", height, "' width='100%' style='max-width:", width, "px; font-family:system-ui,-apple-system,Segoe UI,sans-serif;'>",
    "<text x='12' y='26' font-size='17' font-weight='700' fill='#111827'>MCMC trace plots after burn-in</text>",
    paste(panels, collapse = ""),
    "</svg></div>"
  )
}
```

## Introduction

MaxDiff, also called **Best-Worst Scaling**, is a survey method for measuring relative preference. Instead of asking people to score every item on a 1--10 scale, MaxDiff repeatedly shows a small set of items and asks the respondent to choose the most preferred and least preferred item. In this assignment, the items are MSBA core classes at UCSD Rady, and each valid MaxDiff screen shows four classes.

This approach is useful because people are often better at making trade-offs than they are at assigning calibrated numeric ratings. A respondent may not know whether a class deserves an 8 or a 9, but they can usually decide which class in a small set they most and least prefer. That forced comparison creates clearer information about preference differences and reduces the scale-use bias that can occur with ordinary rating questions.

I analyze the class preference data in three ways. First, I use a simple counts analysis that compares how often each class is selected as best versus worst. Second, I fit a multinomial logit (MNL) model by maximum likelihood. Third, I fit the same MNL model using a Bayesian random-walk Metropolis-Hastings sampler. I expect the three methods to agree on the broad ranking, especially at the top and bottom, while differing slightly in the middle and in how they express uncertainty.

## The Data

```{r read-data}
raw <- read_csv("maxdiff_design_and_choices.csv", show_col_types = FALSE) |>
  rename(
    record_id = `Record ID`,
    task = Task,
    position = Position,
    item = Item,
    response = Response
  )

labels_raw <- read_csv("maxdiff_item_labels.csv", show_col_types = FALSE)
labels <- tibble(
  item = as.integer(labels_raw[[1]]),
  item_label = as.character(labels_raw[[2]])
) |>
  filter(!is.na(item), !is.na(item_label))

raw_task_check <- raw |>
  group_by(record_id, task) |>
  summarize(
    rows_in_task = n(),
    n_best = sum(response == 1),
    n_worst = sum(response == -1),
    all_labeled_items = all(item %in% labels$item),
    .groups = "drop"
  ) |>
  mutate(valid_maxdiff_task = rows_in_task == 4 & n_best == 1 & n_worst == 1 & all_labeled_items)

raw_summary <- tibble(
  statistic = c(
    "Respondents in raw file",
    "Tasks per respondent in raw file",
    "Rows in raw file",
    "Valid 4-item MaxDiff tasks used",
    "Rows used in analysis"
  ),
  value = c(
    n_distinct(raw$record_id),
    paste(unique(raw |> group_by(record_id) |> summarize(n_task = n_distinct(task), .groups = "drop") |> pull(n_task)), collapse = ", "),
    nrow(raw),
    sum(raw_task_check$valid_maxdiff_task),
    sum(raw_task_check$valid_maxdiff_task) * 4
  )
)

raw_summary |> kable()
```

The raw file contains the survey rows plus some follow-up rows that are not part of the 10-class, 4-item MaxDiff design described in the assignment. In particular, some later screens have only two rows and include an unlabeled item number 11. Because the assignment defines the study as 10 class items shown in four-item MaxDiff tasks, I use only tasks that satisfy all of the stated MaxDiff conditions: four rows, one best pick, one worst pick, and only labeled items 1--10.

```{r clean-data}
valid_tasks <- raw_task_check |>
  filter(valid_maxdiff_task) |>
  select(record_id, task)

md <- raw |>
  semi_join(valid_tasks, by = c("record_id", "task"))

task_check <- md |>
  group_by(record_id, task) |>
  summarize(
    rows_in_task = n(),
    n_best = sum(response == 1),
    n_worst = sum(response == -1),
    n_neutral = sum(response == 0),
    .groups = "drop"
  )

basic_descriptives <- tibble(
  statistic = c(
    "Respondents (N)",
    "Tasks per respondent (T)",
    "Items per task (J)",
    "Total labeled items",
    "Total MaxDiff rows used"
  ),
  value = c(
    n_distinct(md$record_id),
    paste(unique(task_check |> count(record_id) |> pull(n)), collapse = ", "),
    paste(unique(task_check$rows_in_task), collapse = ", "),
    n_distinct(labels$item),
    nrow(md)
  )
)

basic_descriptives |> kable()
```

```{r sanity-check}
sanity_summary <- task_check |>
  summarize(
    tasks_checked = n(),
    tasks_with_four_rows = sum(rows_in_task == 4),
    tasks_with_one_best = sum(n_best == 1),
    tasks_with_one_worst = sum(n_worst == 1),
    tasks_with_two_neutral = sum(n_neutral == 2),
    fully_valid_tasks = sum(rows_in_task == 4 & n_best == 1 & n_worst == 1 & n_neutral == 2)
  )

sanity_summary |> kable()
```

The sanity check confirms that every task used in the analysis has exactly one best choice, exactly one worst choice, and two neutral shown-but-not-selected classes.

```{r exposure-counts}
exposure <- md |>
  count(item, name = "times_shown") |>
  left_join(labels, by = "item") |>
  arrange(item)

exposure |> kable()
```

Exposure is reasonably balanced across the 10 labeled classes. This matters because a MaxDiff counts analysis is easiest to interpret when each item has a similar number of opportunities to be chosen.

## Counts Analysis

The counts analysis gives a direct descriptive score for each class. For item \(j\), I compute the percentage of times it was chosen as best and subtract the percentage of times it was chosen as worst:

$$
\text{score}_j = \%\text{best}_j - \%\text{worst}_j.
$$

A score near +1 means the item is almost always selected as best when shown. A score near -1 means it is almost always selected as worst. A score near 0 means the item is either middle-of-the-pack or has mixed reactions.

```{r counts-analysis}
counts_tbl <- md |>
  group_by(item) |>
  summarize(
    times_shown = n(),
    times_best = sum(response == 1),
    times_worst = sum(response == -1),
    pct_best = times_best / times_shown,
    pct_worst = times_worst / times_shown,
    counts_score = pct_best - pct_worst,
    .groups = "drop"
  ) |>
  left_join(labels, by = "item") |>
  arrange(desc(counts_score)) |>
  mutate(counts_rank = row_number())

counts_tbl |>
  transmute(
    item,
    class = item_label,
    times_shown,
    times_best,
    times_worst,
    `% best` = fmt_pct(pct_best),
    `% worst` = fmt_pct(pct_worst),
    score = fmt_num(counts_score)
  ) |>
  kable()
```

```{r bootstrap-counts}
respondents <- unique(md$record_id)

score_for_data <- function(dat) {
  dat |>
    group_by(item) |>
    summarize(
      times_shown = n(),
      times_best = sum(response == 1),
      times_worst = sum(response == -1),
      counts_score = times_best / times_shown - times_worst / times_shown,
      .groups = "drop"
    )
}

boot_scores <- map_dfr(1:1000, function(b) {
  boot_ids <- sample(respondents, length(respondents), replace = TRUE)
  boot_dat <- map_dfr(boot_ids, ~ md |> filter(record_id == .x))
  score_for_data(boot_dat) |>
    mutate(boot = b)
})

count_ci <- boot_scores |>
  group_by(item) |>
  summarize(
    ci_low = quantile(counts_score, 0.025, na.rm = TRUE),
    ci_high = quantile(counts_score, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

counts_tbl <- counts_tbl |>
  left_join(count_ci, by = "item")
```

```{r counts-plot, results='asis'}
knitr::asis_output(svg_bar_chart(
  counts_tbl,
  label = "item_label",
  value = "counts_score",
  low = "ci_low",
  high = "ci_high",
  title = "Counts analysis: best-minus-worst score"
))
```

The counts results show a clear leader: **MGTA 495 Marketing Analytics** has the largest best-minus-worst score by a wide margin. **MGTA 453 Business Analytics** and **MGTA 455 Customer Analytics** are also preferred on this simple metric. The lowest scores are for **MGTA 444 Analytics Consulting**, **MGTA 451 Marketing/Finance/Operations**, and **MGTA 403 AI Math, Programming, & Analytics**. The middle group is much closer together, so small changes in data or modeling approach can move those classes around.

## From MaxDiff Data to MNL Choices

The MaxDiff task can be represented as MNL choice data. On each screen, the respondent first chooses the best item from the full set of four shown classes. If the shown set is \(\{a,b,c,d\}\), the MNL probability that item \(b\) is selected best is

$$
P(\text{best}=b) = \frac{\exp(\beta_b)}{\exp(\beta_a)+\exp(\beta_b)+\exp(\beta_c)+\exp(\beta_d)}.
$$

After the best item is removed, the respondent chooses the worst item from the remaining three. For the worst choice, the utilities are flipped because low utility should make an item more likely to be selected as worst:

$$
P(\text{worst}=c \mid \text{best}=b) =
\frac{\exp(-\beta_c)}{\exp(-\beta_a)+\exp(-\beta_c)+\exp(-\beta_d)}.
$$

Thus, each MaxDiff task contributes two pieces of likelihood information: the best choice from the four-item set and the worst choice from the remaining three-item set. The utility coefficients are identified only up to an additive constant, so I normalize the model by setting \(\beta_1 = 0\) and estimating \(\beta_2, \ldots, \beta_{10}\).

```{r mnl-data}
task_data <- md |>
  arrange(record_id, task, position) |>
  group_by(record_id, task) |>
  summarize(
    items = list(item),
    best = item[response == 1],
    worst = item[response == -1],
    .groups = "drop"
  )

items_mat <- do.call(rbind, task_data$items)
best_vec <- task_data$best
worst_vec <- task_data$worst

row_log_sum_exp <- function(x) {
  m <- apply(x, 1, max)
  m + log(rowSums(exp(x - m)))
}

ll <- function(beta_free) {
  beta <- c(0, beta_free)
  util <- matrix(beta[as.vector(items_mat)], nrow = nrow(items_mat), ncol = ncol(items_mat))

  ll_best <- beta[best_vec] - row_log_sum_exp(util)

  worst_util <- -util
  best_position <- max.col(items_mat == best_vec, ties.method = "first")
  worst_util[cbind(seq_len(nrow(worst_util)), best_position)] <- -Inf
  ll_worst <- -beta[worst_vec] - row_log_sum_exp(worst_util)

  sum(ll_best + ll_worst)
}
```

## MNL via Maximum Likelihood

The log-likelihood sums the probability of the observed best choice and the conditional probability of the observed worst choice across all tasks:

$$
\ell(\boldsymbol{\beta}) = \sum_{\text{tasks}} \left[\beta_{j_b^*} - \log \sum_{j \in \text{shown}} \exp(\beta_j) - \beta_{j_w^*} - \log \sum_{j \in \text{shown}\setminus \{j_b^*\}} \exp(-\beta_j)\right].
$$

```{r mle-fit}
out <- optim(
  par = rep(0, 9),
  fn = ll,
  method = "BFGS",
  hessian = TRUE,
  control = list(fnscale = -1, maxit = 1000)
)

beta_hat <- c(0, out$par)

optim_check <- tibble(
  item = c("optim convergence code", "maximized log-likelihood"),
  value = c(out$convergence, round(out$value, 3))
)

optim_check |> kable()

vcov_free <- solve(-out$hessian)
se_hat <- c(NA, sqrt(diag(vcov_free)))

mle_tbl <- tibble(
  item = 1:10,
  beta_mle = beta_hat,
  se_mle = se_hat,
  mle_share = exp(beta_hat) / sum(exp(beta_hat))
) |>
  left_join(labels, by = "item") |>
  arrange(desc(mle_share)) |>
  mutate(mle_rank = row_number())

mle_tbl |>
  transmute(
    item,
    class = item_label,
    `beta hat` = fmt_num(beta_mle),
    `SE(beta)` = if_else(is.na(se_mle), "reference", as.character(fmt_num(se_mle))),
    `share of preference` = fmt_pct(mle_share),
    rank = mle_rank
  ) |>
  kable()
```

```{r mle-plot, results='asis'}
knitr::asis_output(svg_bar_chart(
  mle_tbl,
  label = "item_label",
  value = "mle_share",
  title = "MNL maximum likelihood shares of preference",
  percent = TRUE
))
```

The MLE ranking is very similar to the counts ranking. **MGTA 495 Marketing Analytics** remains the strongest class by a large margin, followed by **MGTA 453 Business Analytics** and **MGTA 455 Customer Analytics**. The least preferred classes are also similar across methods. The main differences are in the middle, where the counts scores are close. This is expected because the MNL model uses the entire choice structure of each task rather than only summarizing best and worst percentages separately.

## MNL via Bayesian Estimation

For the Bayesian version, I use the same likelihood and combine it with a weakly informative normal prior:

$$
\boldsymbol{\beta} \sim N(\boldsymbol{0}, 10I).
$$

The posterior is proportional to the likelihood times the prior:

$$
\pi(\boldsymbol{\beta} \mid \boldsymbol{y}) \propto L(\boldsymbol{\beta}) \pi(\boldsymbol{\beta}).
$$

I estimate the posterior with a random-walk Metropolis-Hastings sampler. The sampler proposes a new vector \(\boldsymbol{\beta}^*\) by adding normal noise to the current value and accepts or rejects the proposal based on the posterior density ratio.

```{r bayes-fit}
log_post <- function(beta_free) {
  ll(beta_free) + sum(dnorm(beta_free, mean = 0, sd = sqrt(10), log = TRUE))
}

mh <- function(beta0, n_iter, step, keep_draws = TRUE) {
  K <- length(beta0)
  draws <- if (keep_draws) matrix(NA_real_, nrow = n_iter, ncol = K) else NULL
  beta <- beta0
  lp <- log_post(beta)
  accept <- 0

  for (t in seq_len(n_iter)) {
    beta_star <- beta + rnorm(K, mean = 0, sd = step)
    lp_star <- log_post(beta_star)

    if (log(runif(1)) < lp_star - lp) {
      beta <- beta_star
      lp <- lp_star
      accept <- accept + 1
    }

    if (keep_draws) draws[t, ] <- beta
  }

  list(draws = draws, accept_rate = accept / n_iter)
}

# Tune proposal step size with short pilot chains.
candidate_steps <- c(0.03, 0.05, 0.07, 0.10, 0.15)
tuning <- tibble(
  step = candidate_steps,
  accept_rate = map_dbl(candidate_steps, ~ mh(out$par, n_iter = 1500, step = .x, keep_draws = FALSE)$accept_rate)
)

step_use <- tuning$step[which.min(abs(tuning$accept_rate - 0.32))]

mh_out <- mh(out$par, n_iter = 20000, step = step_use, keep_draws = TRUE)
accept_rate <- mh_out$accept_rate
burn <- 5000
post_free <- mh_out$draws[(burn + 1):nrow(mh_out$draws), , drop = FALSE]
post_beta <- cbind(beta_1 = 0, post_free)
colnames(post_beta) <- paste0("beta_", 1:10)

posterior_summary <- tibble(
  item = 1:10,
  beta_post_mean = colMeans(post_beta),
  beta_post_sd = apply(post_beta, 2, sd),
  beta_ci_low = apply(post_beta, 2, quantile, 0.025),
  beta_ci_high = apply(post_beta, 2, quantile, 0.975)
)

post_share_draws <- t(apply(post_beta, 1, function(b) exp(b) / sum(exp(b))))
colnames(post_share_draws) <- paste0("share_", 1:10)

share_summary <- tibble(
  item = 1:10,
  bayes_share = colMeans(post_share_draws),
  share_ci_low = apply(post_share_draws, 2, quantile, 0.025),
  share_ci_high = apply(post_share_draws, 2, quantile, 0.975)
)

bayes_tbl <- posterior_summary |>
  left_join(share_summary, by = "item") |>
  left_join(labels, by = "item") |>
  arrange(desc(bayes_share)) |>
  mutate(bayes_rank = row_number())

tuning |>
  mutate(
    chosen_for_final_chain = step == step_use,
    final_acceptance_rate = if_else(chosen_for_final_chain, accept_rate, NA_real_)
  ) |>
  kable()
```

The final acceptance rate is `r scales::percent(accept_rate, accuracy = 0.1)`, which is in the desired range for a random-walk Metropolis sampler.

```{r trace-plot, results='asis'}
knitr::asis_output(svg_trace_plot(post_beta, burn))
```

The trace plots show stable chains after burn-in rather than a persistent upward or downward drift. This gives a basic visual check that the sampler is exploring a stationary posterior region.

```{r bayes-table}
bayes_tbl |>
  transmute(
    item,
    class = item_label,
    `posterior mean beta` = fmt_num(beta_post_mean),
    `posterior SD beta` = fmt_num(beta_post_sd),
    `beta 2.5%` = fmt_num(beta_ci_low),
    `beta 97.5%` = fmt_num(beta_ci_high),
    `posterior mean share` = fmt_pct(bayes_share),
    rank = bayes_rank
  ) |>
  kable()
```

```{r bayes-vs-mle-se}
mle_tbl |>
  select(item, item_label, beta_mle, se_mle) |>
  left_join(bayes_tbl |> select(item, beta_post_mean, beta_post_sd), by = "item") |>
  filter(item != 1) |>
  arrange(desc(beta_mle)) |>
  transmute(
    item,
    class = item_label,
    `MLE beta` = fmt_num(beta_mle),
    `Bayes mean beta` = fmt_num(beta_post_mean),
    `MLE SE` = fmt_num(se_mle),
    `Bayes posterior SD` = fmt_num(beta_post_sd)
  ) |>
  kable()
```

```{r bayes-plot, results='asis'}
knitr::asis_output(svg_bar_chart(
  bayes_tbl,
  label = "item_label",
  value = "bayes_share",
  low = "share_ci_low",
  high = "share_ci_high",
  title = "Bayesian MNL posterior shares of preference",
  percent = TRUE
))
```

The Bayesian posterior means are close to the MLE estimates because the prior is weak and the sample contains many choices. The posterior standard deviations are also similar to the MLE standard errors for the non-reference coefficients, which is a useful check that the likelihood and sampler are telling a consistent story. The biggest benefit of the Bayesian version is that uncertainty for transformed quantities, such as shares of preference, is easy to summarize: each posterior draw is converted into shares, and the credible intervals come directly from the resulting simulated distribution.

## Comparing the Three Methods

```{r comparison-table}
comparison_tbl <- counts_tbl |>
  select(item, item_label, counts_score, counts_rank) |>
  left_join(mle_tbl |> select(item, mle_share, mle_rank, beta_mle, se_mle), by = "item") |>
  left_join(bayes_tbl |> select(item, bayes_share, bayes_rank, beta_post_mean, beta_post_sd), by = "item") |>
  arrange(mle_rank)

comparison_tbl |>
  transmute(
    item,
    class = item_label,
    `counts score` = fmt_num(counts_score),
    `counts rank` = counts_rank,
    `MLE share` = fmt_pct(mle_share),
    `MLE rank` = mle_rank,
    `Bayes share` = fmt_pct(bayes_share),
    `Bayes rank` = bayes_rank
  ) |>
  kable()
```

```{r comparison-plot, results='asis'}
comparison_chart <- paste0(
  svg_bar_chart(comparison_tbl, "item_label", "counts_score", "Comparison panel 1: Counts score"),
  svg_bar_chart(comparison_tbl, "item_label", "mle_share", "Comparison panel 2: MLE share", percent = TRUE),
  svg_bar_chart(comparison_tbl, "item_label", "bayes_share", "Comparison panel 3: Bayesian share", percent = TRUE)
)
knitr::asis_output(comparison_chart)
```

All three methods agree strongly at the top and bottom of the ranking. **MGTA 495 Marketing Analytics** is the clear top class, while **MGTA 444 Analytics Consulting**, **MGTA 451 Marketing/Finance/Operations**, and **MGTA 403 AI Math, Programming, & Analytics** are near the bottom. The disagreements are mainly in the middle, where several classes have similar estimated preference. That is exactly where sampling noise and model assumptions can flip adjacent ranks.

The MNL model adds structure beyond the counts analysis because it models the choice process explicitly. Every item shown in a task contributes information, including items not selected as best or worst. The Bayesian model adds an especially convenient way to communicate uncertainty for derived quantities such as preference shares. Instead of using a delta method approximation, I can transform every posterior draw into shares and summarize the resulting distribution directly.

## Discussion

Substantively, the survey suggests that students in this sample most prefer **MGTA 495 Marketing Analytics**, followed by classes such as **MGTA 453 Business Analytics** and **MGTA 455 Customer Analytics**. The strength of the preference for MGTA 495 is the clearest result in the analysis: it is ranked first by the counts score, the MLE share, and the Bayesian posterior share. The lower-ranked classes are also fairly stable across methods, although the exact order near the bottom differs slightly.

Methodologically, each method is useful for a different purpose. The counts analysis is simple, transparent, and good for a quick executive summary. The MLE MNL model is better when we want a formal choice model with point estimates, standard errors, and shares of preference. The Bayesian MNL model is most useful when we care about uncertainty in transformed quantities or when we want to extend the model later, for example by estimating individual-level preferences with a hierarchical model.

The main caveat is that this is a self-selected sample of MSBA students, so the results should not be interpreted as the universal ranking of all Rady students. Preferences may also depend on timing, prior coursework, instructor familiarity, and where students are in the program. If I had more time, I would explore heterogeneity by student background and estimate an individual-level or hierarchical model to see whether different groups of students prefer different types of classes.
